Synthesizing Symmetric Lenses

### 2018-07-07 papers/SymmetricLenses/main.tex
blah

\lstset{ language=Caml, basicstyle=\upshape\sffamily,
keywordstyle=\upshape\sffamily\color{dkpurple}, keepspaces=true,
framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
commentstyle=\itshape\rmfamily,
emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disc,merge_left,merge_right,const},
emphstyle=\upshape\sffamily\color{dkpurple}, 
columns=fullflexible,
mathescape, 
xleftmargin=1.5em,
% BCP: I find this distracting:
stringstyle=\sffamily\color{dkblue},
}
\makeatletter
     \let\lst@oldvisiblespace\lst@visiblespace
     \def\lst@visiblespace{\,\lst@oldvisiblespace\,}
\makeatother

@@

\noindent
In a big-data world, one will often find similar data stored in a
variety of different formats. For instance, electronic calendars come in iCalendar, vCalendar and h-event
formats~\cite{calendar-formats}; US taxation data can be transmitted via Tax XML (used by the US government
for e-filing) or TXF (used by TurboTax)~\cite{tax-formats}; and sewing machines can read embroidery files
transmitted as PES formats, DST formats or several others~\cite{embroidery-formats}. \dpw{Editorial comment:
Phil Wadler forced his very irritating long-form citations into the
standard format. ug. how stupid is that.}
\ksf{I like the embroidery example and I also hate the long form citations.}

@@

Lens languages, like any domain-specific language, also cut down the space of
possible programs one can write, and the smaller the space of programs,
the easier it is to find the program one is looking for.  
%This feature can be a negative, if the desired transformation
%falls outside of the scope provided by the language, but can also be positive, as complex
%properties (like invertability) can be enforced automatically.
%Another advantage of reducing the
%set of programs is that doing so can facilitate program synthesis, which is essentially
%a search through the space of possible computations that can be expressed as a lens.
Recent research~\cite{optician, maina+:quotient-synthesis} has exploited this fact by demonstrating that it
is possible to synthesize relatively large, complex lenses from specifications.
In particular, given types that describe source and target formats, there are
synthesis algorithms\bcp{more than one?  I know there are two papers, but
  it's really the same synthesis algorithm} that will find bijective
lenses
% ~\cite{optician}
for files 
in ad-hoc formats, such as those housing linux application configurations.
Oftentimes, these algorithms find such lenses in less than a second, and without
the aid of any examples at all, with just types describing the source and target
formats. 
% \dpw{I would cite the ICFP paper (cite string:
%   maina+:quotient-synthesis) as to appear but I'm not sure how to do that
%   without possibly violating the spirit of double-blind reviewing. ie, if it
%   hasn't appeared yet, how could we know the content if we aren't the authors?
%   Mentioning it in the 3rd person does not seem to suffice ...} \ksf{The ICFP
%   2018 website lists the paper; I think it is okay to cite it. The reviewer can
%   assume we contacted the authors for a copy. We could ask Stephanie for a
%   ruling if we want to be sure. One of the principles of double blind as adopted
%   by SIGPLAN is not to reduce the quality of the paper; omitting this citation
%   would have that affect.}\bcp{+1}\saz{Stephanie says we should cite it in the
%   third person, even if that's a bit awkward, but otherwise citing it is OK.}

@@

The second difficulty with synthesizing symmetric lenses is more
fundamental\dpw{better word?}: While the number of bijective lenses between two
related formats is often tiny, the number of symmetric lenses is often large. In
practice, the difference in cardinality means that when a search algorithm
selects the first bijective lens it finds, that lens has a good chance of being
the one a user wants. In contrast, and to our surprise, our experiments show
that if our search algorithm selects the first symmetric lens it finds, it
typically fails to find the lens the user wants. The flexibity these lenses
admit when it comes to eliding information generates an explosion of possible
transformations. To disambiguate between the many possible lenses, we need new
principles for identifying the ``likely'' ones, and we need new algorithms to
search the space of program transformations using these principles. \dpw{Were
  the bijective lenses chosen using a ``smaller'' criteria? I forget... if not
  edit the prior paragraph.}\afm{no, chosen using a ``first'' criteria.
  Edited.}\dpw{Ideally, we could back up the claim that there are many symmetric
  lenses and not many bijective lenses by graphing the number of such lenses
  somehow. eg, in a past synthesis paper with Steve (POPL 2016) we graphed the
  untyped terms of type $\mathsf{int}\rightarrow\mathsf{int}$. I wonder if we
  can do something similar here.}\afm{What about lenses between ``a''* and
  ''b``*}\bcp{Aren't there an infinite number of those, even if we restrict
  to bijective ones?}

To overcome this second difficulty, we turned to information theory to try to
find a way to characterize the likely and unlikely lenses. Intuitively, a
``likely'' lens is one that propagates a lot of information from source to
target and also from target back to source. Conversely, given the source, the
target is ``unlikely'' to require very much additional information in order to
reconstruct it---if it were\bcp{were what? likely, or unlikely?}, then one is
unlikely to be attempting to synthesize a lens between the two formats\bcp{why?
  even if two big formats only share a little bit of common information,
  mightn't one want a lens between them?}! In order to implement this intuition,
we turn to \emph{stochastic regular expressions} (SREs). Stochastic regular
expressions define a set of strings, as do ordinary regular expressions, but
they also provide the expected distribution of strings within that set. Such
distributions allow us to compute the amount of information gained when
presented an element of the set. For example, the language of the SRE
$\lstinline{"a"} |_{0.5}
\lstinline{"b"}$ contains two elements (the strings
\lstinline{"a"} and
\lstinline{"b"}). The probability 0.5 annotating the union indicates that both
elements are equally likely to appear. Consequently, if we observe an element of
that language (either \lstinline{"a"} or \lstinline{"b"}), we will gain one bit
of information. Said another way, one bit, together with the SRE, suffices to
recover the string in question.\bcp{Is this implying that, if the annotation
  were different, we'd need something other than one bit to recover the
  string??}
\afm{From an information theoretic perspective, yes, because that perspective
  views it as the amount of information if we were communicating a stream of
  these choices.}\bcp{I see---very interesting.  What about mentioning this
  point, to get people ready for some information-theoretic thinking
  later...?} 
On the other hand, consider the language of the SRE \lstinline{c}. It contains 
just one element (the string $c$). If we observe that our algorithm generates
$\lstinline{"c"}$, we gain zero bits of information --- the string was already determined.
Hence\bcp{?}, a lens is ``likely,'' in a technical sense, if relative to the
target SRE $T$, very little information from $t$ is required to generate it.

@@

We implemented this\bcp{which?} algorithm and evaluated its effectiveness on a
range of benchmarks, including data cleaning tasks taken from Flash Fill,
information extraction tasks taken from Augeas, and ad-hoc format
synchronization tasks taken from a variety of locations\footnote{The majority of
  these format synchronization tasks were found from the following index:
  \url{http://fileformats.archiveteam.org/wiki/Electronic_File_Formats}}.

In summary, our main contributions are:
\begin{itemize}
\item We define the space of simple symmetric lenses, which lies between
  asymmetric lenses and full symmetric lenses and has the useful property that
  its members don't have any internal state (\S\ref{sec:overview}). 
\item We show how to synthesize simple symmetric lenses (including all asymmetric lenses as a special case),
using a novel application of stochastic regular expressions to guide the search process.  In particular, we
show that the RE star semi-ring equivalences remain valid when their semantics
are enriched by distributions (\S\ref{sec:synthesis}).
\item We extend the Boomerang implementation with simple symmetric lenses,
  showing that they can easily be combined with a rich variety of other lens
  extensions including quotients and matching lenses. We evaluate our
  implementation over a set of benchmarks drawn from data cleaning, information
  extraction, and file synchronization tasks and demonstrate that we can
  synthesize all of them in under half a minute (\S\ref{sec:evaluation}).
%
  We also analyze the gap between simple and full symmetric lenses
  empirically and argue that simple ones suffice for many applications.
\item We show that the class of simple lenses can be characterized
  semantically as the subset of full symmetric lenses for which a simple
  ``forgetfulness'' property holds (\S\ref{sec:related}).  \bcp{I suggest
    splitting this out into its own section, since there is other related
    work that we'll want to discuss in a Related Work section.}
\end{itemize}

@@

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let salary = number | "N/A"
let emp_salary = name . " " . name . ": " salary
let emp_salaries = "" | emp_salary . ("\n" emp_salary)*
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingC}{\box0}

@@

  \centering
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|>{\columncolor{lightyellow}}c|>{\columncolor{lightgreen}}c|}
      \hline
      \texttt{company.mgmt} data file & \texttt{company.hr} data file \\
      \hline
      \listingA & \listingB \\
      \hline
      \cellcolor{vlightyellow}\texttt{company.mgmt} regular expression & 
      \cellcolor{vlightgreen}\texttt{company.hr}  regular expression\\
      \hline
      \cellcolor{vlightyellow}\listingC & \cellcolor{vlightgreen}\listingD \\
      \hline
    \end{tabular}
  }
  \caption{Hypothetical example data files and corresponding regular expressions
  used by management and HR at a US company to represent employee salaries and health
  insurance providers, respectively. \ksf{Why
  is \lstinline{emp_salaries} written with an or instead of just using
  the star operator? Can we simplify?} }  
  \label{fig:minimized-representations}
\end{figure}

In this section, we give an overview of simple symmetric lenses and
our corresponding synthesis algorithm using an example drawn from a
hypothetical U.S. company.  In this company, management and
human resources (HR) store information about employees in separate files:
management records the names of employees and their salaries while
HR stores the names of employees and their health
insurance providers.  Figure~\ref{fig:minimized-representations} gives
examples of the two types of files and regular expressions describing
the two formats.
The company uses simple symmetric lenses to keep these files synchronized.
If management adds a new employee, ``Chris Roe: 32500'',  a lens adds 
the corresponding entry ``Chris Roe, UNK'' to HR's file.  The
sentinel value ``UNK'' represents the fact that the employee's
insurance company is currently unknown.  A similar update happens if
HR adds a new employee before management, in which case the sentinel
value ``N/A'' reflects the unknown salary information\bcp{can we just use
  ``UNKNOWN'' on both sides?  Or is there some reason to make it more
  complicated?} 
Furthermore, if HR
corrects an error in an employee's name, say changing ``John Public''
to ``Jon Public'', the lens mirrors that name change in management's
file. Not all the data is mirrored, however. The management file 
is not updated in response to insurance changes, and the HR file
is oblivious to salary changes.  Simple symmetric lenses are
appropriate for keeping these two files synchronized because the files have
a mix of shared and unshared information.

@@

\subsection{Simple Symmetric Lenses}
A simple symmetric lens between formats described by types $\Regex$
and $\RegexAlt$ 
is defined by four functions\bcp{I think a picture would be helpful.  I can probably
  make one pretty easily by tweaking the one from the original symmetric paper.}

@@

\noindent
that collectively satisfy four ``round-tripping'' laws: 
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{s})}{s} = s\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{t})}{t} = t\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{s}{t})}{s} = s\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{t}{s})}{t} = t
\end{gather}
\noindent
The two create functions provide default values for times when
data appears on one side but not the other (\EG, the ``N/A'' salary
entry added to the management file when HR inserts a new employee).
The two put functions propagate edits from one format to the
other by combining the change from one file with the previous value
from the other file. The roundtripping laws guarantee that identify edits to one file do
not change the data in the other file.

@@

The simplest combinator is the identity lens \IdentityLens, which takes as an
argument a regular expression $\Regex$ and performs an identity mapping on data
matching $\Regex$. The identity lens fully connects the
corresponding data in the two formats: when one format is
updated, the other format gets the exact same value. In our example,
when a \lstinline{name} is updated in one file, the identify lens ensures it is
also updated in the other file.

@@

In contrast to the identity lens, the disconnect lens, \Disconnect, leaves data
unconnected.\bcp{Can we call it ``disconnect''?} The disconnect lens takes four arguments: two regular 
expressions and two strings. The regular expressions specify the
formats of the corresponding data on the two sides while the strings provide default
values.  In our extended example, 
the \lstinline{salary} and \lstinline{company} fields are independent,
so we use the disconnect lens to relate them.
The defaults are strings that indicate the information is missing.
%
\begin{lstlisting}
disc(salary, company, "N/A", "UNK") : salary $\Leftrightarrow$ company
\end{lstlisting}
%
The insert \lstinline{ins} and delete \lstinline{del} lenses are
syntactic sugar for uses of the disconnect lens in which a string
constant is omitted entirely from the source and target formats,
respectively. \ksf{Are these in the right order?}
\begin{lstlisting}
 ins($t$) = disc("", $t$, "", $t$)
 del($s$) = disc($s$, "", $s$, "")
\end{lstlisting}

Finally, there are a number of standard lenses that compositionally
construct more complex lenses from simpler ones.  These combinators include 
\lstinline{concat} for concatenation, 
\lstinline{or} for variation, and
\IterateLens for sequences.
We usually write \lstinline{concat} using an infix
notation \lstinline{$\Lens_1$ . $\Lens_2$}
as in:
%
\begin{lstlisting}
id(name) . ins(",") . id(name) . del(": ") . ins(",")
\end{lstlisting}
% concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
%
which appears in the definition of the \lstinline{name_lens} in our
running example. 
%\ksf{I don't think this example is necessary.}
%Data can also appear in multiple forms. For example,
%a list of employees with insurance information can either be empty or
%be a sequence of employee/insurance records: \ksf{This example is
%vacuous as you could eliminate the or in favor of the *.  Do somethign
%else to introduce or?}
%
%\begin{lstlisting}
%or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
%\end{lstlisting}
The \IterateLens lens also appears in our running example, where it
synchronizes data for a list of employees when given a
lens \lstinline{employee_lens} that synchronizes  data for a single employee:
%
\begin{lstlisting}
iterate(id("\n") . employee_lens) : ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_ins)$^*$
\end{lstlisting}
%

\begin{figure}
\begin{lstlisting}
let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
let employee_lens = name_lens . disc(salary,company,"N/A","UNK") in
let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
ins(header) . employees_lens : emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
%\ksf{I don't think this is necesssary.}
%let emp_lens = or(id(""),employees_lens) in
  \caption{A lens, written in our language for building simple symmetric lenses
    on string data, that synchronizes management and HR employee files.
    Note that we can annotate our lenses with types, to confirm the lens
    operates on the languages of the regular expressions. In this work, we
    synthesize the lens itself from such regular expressions.}
  \label{fig:example_lens}
\end{figure}

@@

We provide typing rules for each of the simple symmetric combinators,
most of which are syntax directed.  There is one additional
typing rule, which allows us to use equivalence relations on regular
expressions to convert the type of any lens to a semantically equal
but syntactically different pair of regular expressions.  The flexibility this
type equivalence rule provides is useful when a different
representation of a regular expression would facilitate synthesis.

@@

Note that, while the synchronization task for our running example is quite simple, the
requisite lens is relatively complex. Furthermore, this complexity is
inherent in manually writing simple symmetric lenses because the
lenses must precisely describe the portions of the data that should be
connected and those that should be disconnected.  For disconnected
pieces, the lens must further specifiy appropriate defaults.

@@

To alleviate the difficulty of manually writing simple symmetric
lenses, we wish to synthesize them instead. Given 
a pair of regular expressions and a set of input-output examples, we want to
find a simple symmetric lens that is typed by the regular expression
pair and satisfies the input/output examples. We call such a lens
a \emph{satisfying lens}.
In our running example, we wish to
synthesize a lens between \lstinline{emp_salaries} and \lstinline{emp_insurance},
using as an input-output example the data in
Figure~\ref{fig:minimized-representations}.  The algorithm uses such
examples to find suitable defaults for \Disconnect lenses and to
figure out how to match data that could be synchronised in multiple
ways.

A challenge is that the simple symmetric lens combinators permit many
well-typed lenses
between a given pair of regular expressions. For example, 
Figure~\ref{fig:example_lens} gives one possible lens between regular
expressions  \lstinline{emp_salaries} and \lstinline{emp_insurance},
but 
%
\begin{lstlisting}
disc(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
\end{lstlisting}
%
is another valid lens (where \lstinline{rep($\Regex$)} is a string in the
language of $\Regex$). Furthermore, adding the input-output example in
Figure~\ref{fig:minimized-representations} 
helps, but only a little, because it requires many examples to rule
out all the possible occurrences of nested \Disconnect lenses,
particularly in complex formats. 
%\ksf{I don't understand the following.  We shouldn't explain the issue in
%terms of \PutR and \PutL as we are focused on the combinators rather
%than the raw functions.}
%The default strings given to \Disconnect merely need to be the strings
%in the input-output examples, and the lens will still satisfy the
%specification.  While the disconnect lens is prohibited by two
%examples, or by a carefully chosen input-output example on \PutR
%or \PutL, this problem continues presenting itself for the sublenses,
%and is highlighted even more on larger formats.
While
we could ask programmers to provide detailed and precise specifications for their
synthesis tasks, we would prefer the algorithm to choose the right lens
automatically. Instead of merely finding \emph{any} satisfying lens, we wish to
synthesize a satisfying lens that is likely to please the user.

@@

In our setting, we already have a way of expressing sets of data:
regular expressions.  What we need is a way to express probability
distributions over that data.
To that end, we adopt \emph{stochastic regular expressions}\ksf{is
there a citation?}, 
which are regular expressions where each operator is annotated with a
probability.  Stochastic regular expressions
allow us to jointly specify the allowable data and give a probability
distribution over it.  From this information, we can calculate the
entropy of the stochastic regular expression.

>>>>

\lstset{ language=Caml, basicstyle=\upshape\sffamily,
keywordstyle=\upshape\sffamily\color{dkpurple}, keepspaces=true,
framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
commentstyle=\itshape\rmfamily,
emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disconnect,merge_left,merge_right,const},
emphstyle=\upshape\sffamily\color{dkpurple}, 
columns=fullflexible,
mathescape, 
xleftmargin=1.5em,
% BCP: I find this distracting:
stringstyle=\sffamily\color{dkblue},
}
\makeatletter
     \let\lst@oldvisiblespace\lst@visiblespace
     \def\lst@visiblespace{\,\lst@oldvisiblespace\,}
\makeatother

@@

\noindent
In a big-data world, one will often find similar data stored in a variety of
different formats. For instance, electronic calendars come in iCalendar,
vCalendar and h-event formats~\cite{calendar-formats}; US taxation data can be
transmitted via Tax XML (used by the US government for e-filing) or TXF (used by
TurboTax)~\cite{tax-formats}; and sewing machines can read embroidery files
transmitted as PES formats, DST formats or several
others~\cite{embroidery-formats}.

@@

Lens languages, like any domain-specific language, also cut down the space of
possible programs one can write, and the smaller the space of programs,
the easier it is to find the program one is looking for.  
%This feature can be a negative, if the desired transformation
%falls outside of the scope provided by the language, but can also be positive, as complex
%properties (like invertability) can be enforced automatically.
%Another advantage of reducing the
%set of programs is that doing so can facilitate program synthesis, which is essentially
%a search through the space of possible computations that can be expressed as a lens.
Recent research~\cite{optician, maina+:quotient-synthesis} has exploited this
fact by demonstrating that it is possible to synthesize relatively large,
complex lenses from specifications. In particular, given types that describe
source and target formats, the bijective synthesis algorithm presented in this
research will find bijective lenses
% ~\cite{optician}
for files in ad-hoc formats, such as those housing linux application
configurations. Oftentimes, these algorithms find such lenses in less than a
second, and without the aid of any examples at all, with just types describing
the source and target formats.
% \dpw{I would cite the ICFP paper (cite string:
%   maina+:quotient-synthesis) as to appear but I'm not sure how to do that
%   without possibly violating the spirit of double-blind reviewing. ie, if it
%   hasn't appeared yet, how could we know the content if we aren't the authors?
%   Mentioning it in the 3rd person does not seem to suffice ...} \ksf{The ICFP
%   2018 website lists the paper; I think it is okay to cite it. The reviewer can
%   assume we contacted the authors for a copy. We could ask Stephanie for a
%   ruling if we want to be sure. One of the principles of double blind as adopted
%   by SIGPLAN is not to reduce the quality of the paper; omitting this citation
%   would have that affect.}\bcp{+1}\saz{Stephanie says we should cite it in the
%   third person, even if that's a bit awkward, but otherwise citing it is OK.}

@@

The second difficulty with synthesizing symmetric lenses is more
fundamental\dpw{better word?}: While the number of bijective lenses between two
related formats is often tiny, the number of symmetric lenses is often large. In
practice, the difference in cardinality means that when a search algorithm
selects the first bijective lens it finds, that lens has a good chance of being
the one a user wants. In contrast, and to our surprise, our experiments show
that if our search algorithm selects the first symmetric lens it finds, it
typically fails to find the lens the user wants. The flexibity these lenses
admit when it comes to eliding information generates an explosion of possible
transformations. To disambiguate between the many possible lenses, we need new
principles for identifying the ``likely'' ones, and we need new algorithms to
search the space of program transformations using these principles.

To overcome this second difficulty, we turned to information theory to try to
find a way to characterize the likely and unlikely lenses. Intuitively, a
``likely'' lens is one that propagates a lot of information from source to
target and also from target back to source. Conversely, given the source, the
target is ``unlikely'' to require very much additional information in order to
reconstruct it, under the assumption that users typically want to synchronize
formats with a large amount of shared information. In order to implement this
intuition, we turn to \emph{stochastic regular expressions} (SREs). Stochastic
regular expressions define a set of strings, as do ordinary regular expressions,
but they also provide the expected distribution of strings within that set. Such
distributions allow us to compute the amount of information gained when
presented an element of the set. For example, the language of the SRE
$\lstinline{"a"} |_{0.5}
\lstinline{"b"}$ contains two elements (the strings \lstinline{"a"} and \lstinline{"b"}).
The probability 0.5 annotating the union indicates that both elements are
equally likely to appear. Consequently, if we observe an element of that
language (either \lstinline{"a"} or
\lstinline{"b"}), we will gain one bit of information. Said another way, one bit, together with
the SRE, suffices to recover the string in question. Interestingly, if the
probability annotation were anything but 0.5, the expected information content
(or entropy) would actually be less than 1 (given a stream of such choices, on
average we could encode the the contents of stream in fewer bits than the the
number of choices present). On the other hand, consider the language of the SRE
\lstinline{c}. It contains just one element (the string $c$). If we observe that
our algorithm generates $\lstinline{"c"}$, we gain zero bits of information ---
the string was already determined. Using this foundation, a lens $\Lens \OfType
\Regex \Leftrightarrow \RegexAlt$ is ``likely,''
in a technical sense, if given a string $s$ in the source $S$, very little information
is needed to determine which synchronized string $t$ is present in the target
$T$.

@@

We implemented the \GreedySynth and \RXSearch algorithms and evaluated their
effectiveness of synthesizing synchronization functions on a range of
benchmarks, including data cleaning tasks taken from Flash Fill, information
extraction tasks taken from Augeas, and ad-hoc format synchronization tasks
taken from a variety of locations\footnote{The majority of these format
  synchronization tasks were found from the following index:
  \url{http://fileformats.archiveteam.org/wiki/Electronic_File_Formats}}.

In summary, our main contributions are:
\begin{itemize}
\item We define the space of simple symmetric lenses, which lies between
  asymmetric lenses and full symmetric lenses and has the useful property that
  its members don't have any internal state (\S\ref{sec:overview}). 
\item We show how to synthesize simple symmetric lenses (including all asymmetric lenses as a special case),
using a novel application of stochastic regular expressions to guide the search process.  In particular, we
show that the RE star semi-ring equivalences remain valid when their semantics
are enriched by distributions (\S\ref{sec:synthesis}).
\item We extend the Boomerang implementation with simple symmetric lenses,
  showing that they can easily be combined with a rich variety of other lens
  extensions including quotients and matching lenses. We evaluate our
  implementation over a set of benchmarks drawn from data cleaning, information
  extraction, and file synchronization tasks and demonstrate that we can
  synthesize all of them in under half a minute (\S\ref{sec:evaluation}).
%
  We also analyze the gap between simple and full symmetric lenses
  empirically and argue that simple ones suffice for many applications.
\item We show that the class of simple lenses can be characterized semantically
  as the subset of full symmetric lenses for which a simple ``forgetfulness''
  property holds (\S\ref{sec:relationship}).
\end{itemize}

@@

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let salary = number | "unk"
let emp_salary = name . " " . name . ": " salary
let emp_salaries = "" | emp_salary . ("\n" emp_salary)*
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingC}{\box0}

@@

  \centering
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|>{\columncolor{vlightyellow}}c|>{\columncolor{vlightgreen}}c|}
      \hline
      \texttt{company.mgmt} data file & \texttt{company.hr} data file \\
      \hline
      \listingA & \listingB \\
      \hline
      \cellcolor{lightyellow}\texttt{company.mgmt} regular expression & 
      \cellcolor{lightgreen}\texttt{company.hr}  regular expression\\
      \hline
      \cellcolor{lightyellow}\listingC & \cellcolor{lightgreen}\listingD \\
      \hline
    \end{tabular}
  }
  \caption{Hypothetical example data files and corresponding regular expressions
  used by management and HR at a US company to represent employee salaries and health
  insurance providers, respectively. \ksf{Why
  is \lstinline{emp_salaries} written with an or instead of just using
  the star operator? Can we simplify?}\afm{It does require the or operator, if
  we don't want it to have trailing or preceding newlines.  Howevever, I
  intentionally had this as not having those to force a discussion about
  rewrites and searching through rewrites.} }  
  \label{fig:minimized-representations}
\end{figure}

In this section, we give an overview of simple symmetric lenses and our
corresponding synthesis algorithm using an example drawn from a hypothetical
U.S. company. In this company, management and human resources (HR) store
information about employees in separate files: management records the names of
employees and their salaries while HR stores the names of employees and their
health insurance providers. Figure~\ref{fig:minimized-representations} gives
examples of the two types of files and regular expressions describing the two
formats. The company uses simple symmetric lenses to keep these files
synchronized. If management adds a new employee, ``Chris Roe: 32500'', a lens
adds the corresponding entry ``Chris Roe, UNK'' to HR's file. The sentinel value
``UNK'' represents the fact that the employee's insurance company is currently
unknown. A similar update happens if HR adds a new employee before management,
in which case the sentinel value ``unk'' reflects the unknown salary information
Furthermore, if HR corrects an error in an employee's name, say changing ``John
Public'' to ``Jon Public'', the lens mirrors that name change in management's
file. Not all the data is mirrored, however. The management file is not updated
in response to insurance changes, and the HR file is oblivious to salary
changes. Simple symmetric lenses are appropriate for keeping these two files
synchronized because the files have a mix of shared and unshared information.

@@

\subsection{Simple Symmetric Lenses}
A simple symmetric lens between formats described by types $\Regex$
and $\RegexAlt$ 
is defined by four functions\bcp{I think a picture would be helpful.  I can probably
  make one pretty easily by tweaking the one from the original symmetric paper.}
\afm{That would be awesome.}

@@

\noindent
that collectively satisfy four ``round-tripping'' laws: 
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{s})}{s} = s\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{t})}{t} = t\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{s}{t})}{s} = s\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{t}{s})}{t} = t
\end{gather}
\noindent
The two create functions provide default values for times when
data appears on one side but not the other (\EG, the ``unk'' salary
entry added to the management file when HR inserts a new employee).
The two put functions propagate edits from one format to the
other by combining the change from one file with the previous value
from the other file. The roundtripping laws guarantee that identify edits to one file do
not change the data in the other file.

@@

The simplest combinator is the identity lens \IdentityLens, which takes as an
argument a regular expression $\Regex$ and performs an identity mapping on data
matching $\Regex$. The identity lens fully connects the
corresponding data in the two formats: when one format is
updated, the other format gets the exact same value. In our example,
when a \lstinline{name} is updated in one file, the identify lens ensures it is
also updated in the other file. In this manner, both the \CreateR and \CreateL
functions are the identity function ($\CreateR \App s = s$), and the put
functions merely return the first argument ($\PutROf{s_1}{s_2} = s_2$).

@@

In contrast to the identity lens, the disconnect lens, \Disconnect, leaves data
unconnected. The disconnect lens takes four arguments: two regular expressions
and two strings. The regular expressions specify the formats of the
corresponding data on the two sides while the strings provide default values.
On creates, the input values are thrown away, and default values are returned ($\CreateROf{\String} =
\lstinline{"UNK"}$), and on puts, the second argument is used, and the first is
thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$).  In
our extended example, the \lstinline{salary} and
\lstinline{company} fields are independent, so we use the disconnect lens to relate them. The defaults are strings that indicate the information is missing.
%
\begin{lstlisting}
disconnect(salary, company, "unk", "UNK") : salary $\Leftrightarrow$ company
\end{lstlisting}
%
The insert \lstinline{ins} and delete \lstinline{del} lenses are
syntactic sugar for uses of the disconnect lens in which a string
constant is omitted entirely from the source and target formats,
respectively.
\begin{lstlisting}
 ins($t$) = disconnect("", $t$, "", $t$)
 del($s$) = disconnect($s$, "", $s$, "")
\end{lstlisting}

Finally, there are a number of standard lenses that compositionally construct
more complex lenses from simpler ones. These combinators include concatenation
($\ConcatLensOf{\Lens_1}{\Lens_2}$ or $\Lens_1 . \Lens_2$), variation
($\OrLensOf{\Lens_1}{\Lens_2}$), and iteration ($\IterateLensOf{\Lens}$).  For example;
%
\begin{lstlisting}
id(name) . ins(",") . id(name) . del(": ") . ins(",")
\end{lstlisting}
% concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
%
appears in the definition of the \lstinline{name_lens} in our running example.
%\ksf{I don't think this example is necessary.}
%Data can also appear in multiple forms. For example,
%a list of employees with insurance information can either be empty or
%be a sequence of employee/insurance records: \ksf{This example is
%vacuous as you could eliminate the or in favor of the *.  Do somethign
%else to introduce or?}
%
%\begin{lstlisting}
%or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
%\end{lstlisting}
The \IterateLens lens also appears in our running example, where it
synchronizes data for a list of employees when given a
lens \lstinline{employee_lens} that synchronizes  data for a single employee:
%
\begin{lstlisting}
iterate(id("\n") . employee_lens) : ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_ins)$^*$
\end{lstlisting}
%

\begin{figure}
\begin{lstlisting}
let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
let employee_lens = name_lens . disconnect(salary,company,"unk","UNK") in
let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
ins(header) . employees_lens : emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
%\ksf{I don't think this is necesssary.}
%let emp_lens = or(id(""),employees_lens) in
  \caption{A lens, written in our language for building simple symmetric lenses
    on string data, that synchronizes management and HR employee files.
    Note that we can annotate our lenses with types, to confirm the lens
    operates on the languages of the regular expressions. In this work, we
    synthesize the lens itself from such regular expressions.}
  \label{fig:example_lens}
\end{figure}

@@

We provide typing rules for each of the simple symmetric combinators, most of
which are syntax directed. However, regular expressions have a number of
equivalences on them, and sometimes we must convert one regular expression type
to another, equivalent regular expression type. Our typing rules include type
equivalence, so if a lens is typed between two regular expressions, it is also
typed between two equivalent ones. The flexibility this type equivalence rule
provides is useful when a different representation of a regular expression would
facilitate synthesis.

@@

@@

While our example is relatively simple for demonstration, writing such lenses
for real-world formats is quite difficult, as these formats can be large and
transform data in complex and unintuitive ways. To alleviate the difficulty of
manually writing simple symmetric lenses, we wish to synthesize them instead.
Given a pair of regular expressions and a set of input-output examples, we want
to find a simple symmetric lens that is typed by the regular expression pair and
satisfies the input/output examples. We call such a lens a \emph{satisfying
  lens}. In our running example, we wish to synthesize a lens between
\lstinline{emp_salaries} and \lstinline{emp_insurance}, using as an input-output
example the data in Figure~\ref{fig:minimized-representations}. The algorithm
uses such examples to find suitable defaults for \Disconnect lenses and to
figure out how to match data that could be synchronised in multiple ways.

A challenge is that the simple symmetric lens combinators permit many
well-typed lenses
between a given pair of regular expressions. For example, 
Figure~\ref{fig:example_lens} gives one possible lens between regular
expressions  \lstinline{emp_salaries} and \lstinline{emp_insurance},
but 
%
\begin{lstlisting}
disconnect(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
\end{lstlisting}
%
is another valid lens (where \lstinline{rep($\Regex$)} is a string in the
language of $\Regex$). Furthermore, adding the input-output example in
Figure~\ref{fig:minimized-representations} 
helps, but only a little, because it requires many examples to rule
out all the possible occurrences of nested \Disconnect lenses,
particularly in complex formats. 
%\ksf{I don't understand the following.  We shouldn't explain the issue in
%terms of \PutR and \PutL as we are focused on the combinators rather
%than the raw functions.}
%The default strings given to \Disconnect merely need to be the strings
%in the input-output examples, and the lens will still satisfy the
%specification.  While the disconnect lens is prohibited by two
%examples, or by a carefully chosen input-output example on \PutR
%or \PutL, this problem continues presenting itself for the sublenses,
%and is highlighted even more on larger formats.
While we could ask programmers to provide detailed and precise specifications
for their synthesis tasks, we would prefer an algorithm that can intelligently
choose a lens despite the underspecification. Instead of merely finding
\emph{any} satisfying lens, we wish to synthesize a satisfying lens that is
likely to please the user.

@@

In our setting, we already have a way of expressing sets of data:
regular expressions.  What we need is a way to express probability
distributions over that data.
To that end, we adopt \emph{stochastic regular expressions}\cite{?}, 
which are regular expressions where each operator is annotated with a
probability.  Stochastic regular expressions
allow us to jointly specify the allowable data and give a probability
distribution over it.  From this information, we can calculate the
entropy of the stochastic regular expression.
